---
title: "PML_Course"
author: "Joseph Raj A"
date: "Sunday, June 21, 2015"
output: html_document
---
Installed required packages like Caret, randomForest and rpart
install.packages("caret")
install.packages("randomForest")
install.packages("rpart")  

```{r}
library(caret)
library(randomForest)
library(rpart)
```

Overall seed for reproduceability has been set
```{r}
set.seed(1234)
```

Accessing the given URL and saving both datasets as training and testing.
```{r}
training_Url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testing_Url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
training <- read.csv(url(training_Url), na.strings=c("NA","DIV/0!",""))
testing <- read.csv(url(testing_Url), na.strings=c("NA","DIV/0!",""))
```

After Analyzing the data, Columns with values as string "NA" / "DIV/0!" / "" are not required for sampling and hence,these will be changed to NA and then deleted as its not required.
```{r}
training <-training[,colSums(is.na(training)) == 0]
testing <-testing[,colSums(is.na(testing)) == 0]
```

The First 7 columns user_name, raw_timestamp_part_1, raw_timestamp_part_2, cvtd_timestamp, new_window and num_window which seem to be irrelevant variables for this project, hence deleted all these variables (1 to 7 columns)
```{r}
training <- training[,-c(1:7)]
testing <- testing[,-c(1:7)]
```

Checking the dataset
```{r}
dim(training)
dim(testing)
```

Partioning the data 80% training 20% testing and lets plot to see how the data behaves.
```{r}
samples <- createDataPartition(y=training$classe, p=0.80, list=FALSE)
trainingset <- training[samples, ] 
testingset <- training[-samples, ]
```

Checking the sample datasets
```{r}
dim(trainingset)
dim(testingset)
```

Applying the random Forest Algorithm on training dataset
```{r}
rfmodel <- randomForest(classe ~. , data=trainingset, method="class")
```

Based on the built model, predicting the testing dataset
```{r}
prediction <- predict(rfmodel, testingset, type = "class")
```

Running the ConfusionMatrix on the prediction to see the results.
```{r}
confusionMatrix(prediction, testingset$classe)
```

#Decision

##Accuracy for Random Forest model was 0.998 (95% CI: (0.996, 0.9991)). The expected out-of-sample error is estimated at 0.002, or 0.2%. The expected out-of-sample error is calculated as 1 - accuracy for predictions made against the cross-validation set. 

##Our Test data set comprises 20 cases. With an accuracy above 99% on our cross-validation data, we can expect that very few, or none, of the test samples will be missclassified

##Running the prediction model on original Testing dataset using Random Forest algorithm to get the desired outputs for submission.

```{r}
finaloutput <- predict(rfmodel, testing, type="class")
finalOutput
```

#Preparing the files for submission
```{r}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
```